\documentclass{article}
\usepackage[layout=letterpaper,margin=1in]{geometry}

\usepackage{fancyhdr}
\usepackage{xspace}
\usepackage{color}
\usepackage{comment}

\usepackage[colorlinks=true]{hyperref}

\usepackage[all=normal,lists]{savetrees}

\usepackage[endianness=big]{bytefield}
\bytefieldsetup{boxformatting={\centering\footnotesize}}

\usepackage{etoc}
\setcounter{tocdepth}{4}

\newcommand{\colorbitbox}[3]{%
\rlap{\bitbox{#2}{\color{#1}\rule{\width}{\height}}}%
\bitbox{#2}{#3}}
\definecolor{lightblue}{RGB}{0,204,255}
\definecolor{lightcyan}{rgb}{0.84,1,1}
\definecolor{lightgreen}{rgb}{0.64,1,0.71}
\definecolor{lightergreen}{rgb}{0.84,1,0.87}
\definecolor{lightred}{rgb}{1,0.7,0.71}

\usepackage{enumitem}
\setlistdepth{8}
%\setlist[itemize,1]{label=$\bullet$}
\setlist[itemize,1]{label=}
\setlist[itemize,2]{label=--}
\setlist[itemize,3]{label=$\circ$}
\setlist[itemize,4]{label=$\rightarrow$}
\setlist[itemize,5]{label=$\diamondsuit$}
\setlist[itemize,6]{label=$\cdot$}
\setlist[itemize,7]{label=$\star$}
\setlist[itemize,8]{label=$\ast$}
\renewlist{itemize}{itemize}{8}

\begin{document}

\fancyfoot[L]{Revision 0.2.4 -- Dec 20, 2012}
\pagestyle{fancyplain}

\title{ICE Debug Board FPGA Interface protocol}
\author{Ben Kempke, Pat Pannuto \{bpkempke,ppannuto\}@umich.edu}
\date{}
\maketitle

\section*{Introduction}
The ICE board is a dual-purpose development board, enabling both high-speed
programming (via DMA) of M3 chips and interfacing with live M3 sessions and
external peripherals.

This means that libraries must be able to gracefully handle asychronous,
unexpected events. In particular, some semantics resemebling transactions
would be desirable, as well as an effort to maintain event ordering over the
serial communications link. The consequence is that this document should be
viewed less as a protocol specification and more of a living document as we
explore the best semantics for this domain.

At a high-level, every message is a composition of an {\em event id}, a
{\em message type} and an optional {\em message}. Every message sent {\em to}
the ICE board {\bf MUST} be replied to by either an ACK or NAK response.
Every message sent {\em from} the ICE board {\bf MUST
NOT} be acknowledged. The rationale for this asymmetry is to simplify the ICE
hardware design. With this mechanism, the hardware is not obligated to
preserve any communication state beyond the immediate running context. The
event ids provide a total ordering of the event history, insofar as is
possible.

\setcounter{tocdepth}{2}
%http://tex.stackexchange.com/questions/87709/minitoc-of-a-subsection/87716
%\etocsettocstyle{\subsection*{\contentsname}\hrule\medskip
%\begin{minipage}{.95\linewidth}}
%{\end{minipage}\medskip\hrule}
%\etocsettocstyle
%    {\subsection*{\contentsname}\hrule\medskip
%        \everypar{\rightskip.1\linewidth}}
%    {\nobreak\medskip\hrule\bigskip}
\tableofcontents

\clearpage
\section{Event Ids}
\label{sec:ids}

\subsection{Motivation}
Some decisions in microcontrollers necessarily race. As contrived example, if
two GPIO pins are defined as interrupts but are electrically connected in the
external circuit, and then line is pulled high, which interrupt fires first?
While the decision is arbitrary, there is motivation to define an ordering of
events in the system. The events observed by ICE can be replayed in the
M-ultaor for debugging, but the is only possible if they can be accurately
re-created.

In practice, this diverges to two ideas:

\subsubsection{Concurrent Events}
Two events could be labeled with the same event id, indicating that they
occurred too close together in time for the ICE to distinguish them. This
would require divergence in the simulator for debugging runs; do-able, but not
trivial.

\subsubsection{I/O Pass-Through}
\label{sec:pass-through}
Currently, the FPGA and the M3 share GPIO pins (Nx2 header):
\begin{verbatim}
          /-----\
FPGA ---- | o o | ----- M3
          \-----/
          GPIO N
\end{verbatim}
Instead perhaps we can explicitly pass all I/O through the FPGA:
\begin{verbatim}
/---\
| o | ---- FPGA -- {TP} -- M3
\---/
GPIO N
\end{verbatim}
This has the disadvantage of doubling the number of FPGA I/O pins consumed for
each of the M3 I/O pins. It does enable the FPGA to strictly define an order
of events. We can possibly explore this with the current board by mapping
GPIOs 16-23 as FPGA-input only, jumpering over GPIOs 8-15 and having the FPGA
drive those, leaving GPIOs 0-7 as the first one.

\subsection{Implementation}

Event IDs are an unsigned single byte number. They define a total ordering the
events actually occurred in the system. In particular, if a command message is
being sent to set GPIO 0 (an output) high at the same time that GPIO 1 (an
input) goes high, the ordering will be {\tt Event N: GPIO 1 --> High} then
{\tt Event N+1: GPIO 0 --> High}. With \ref{sec:pass-through}'s pass through,
perhaps the GPIO 1 transition could be delayed during command message
reception, but this is for further exploration.

For message format consistency, event ids are included in both directions of
communication, however the field may be safely ignored by the FPGA. The FPGA
itself must by definition have some order of I/O events it processed, which
are encapsulated by these event ids. The event id of a control message is
assigned by the FPGA whenever it actually processes the event and is indicated
to the controller via the ACK message.

\clearpage
\section{Messages / Base Protocol}
\label{sec:messages}

Messages are composed of a
one-byte message type identifier
followed by a one-byte event identifier
followed by a one-byte unsigned length indicator
followed by an optional message component. Visually:
\begin{quote}
\begin{bytefield}{32} \\
\colorbitbox{lightgreen}{8}{Message Type} &
\colorbitbox{lightred}{8}{Event ID} &
\colorbitbox{lightcyan}{8}{Length} &
\bitbox{8}{[Message Data...]} &
\end{bytefield}
\end{quote}
Effort
should be made to keep type identifiers within the ASCII range where
reasonable, mapped to appropriate letters.

There are two types of transactions defined: {\em synchronous} and {\em
asynchronous}. A synchronous message is one initiated by the controlling PC
and must be responded to by a \{N\}ACK from the ICE board. An asychronous
message is generated by the ICE board in response to a hardware event. Only
one synchronous message is permitted to be live at any given time. The ICE
board may send any number of asychronous messages before responding to the
synchronous message.

The protocol in use is undefined
until a version request is ACKed. The reception of a {\em NAK with reason}
with a preferred version is {\bf NOT} sufficient to establish a version, the
controller {\bf MUST} explicilty send another {\bf v}ersion request and
receive an ACK to establish the protocol version in use.

{\bf Only `V' and `v' messages are valid until a version has been negotiated.}
Any asynchronous hardware events may either be queued or discarded.

\medskip

The base protocol defines the following immutable message types and their
properties:

\begin{itemize}

\item {\textbf {\texttt 0x00} ACK}
\item {\textbf {\texttt 0x01} NAK}
\item {\em Synchronous Response}
\begin{itemize}
	\item[]
\begin{bytefield}{32} \\
\colorbitbox{lightgreen}{8}{0x00 or 0x01} &
\colorbitbox{lightred}{8}{Event ID} &
\colorbitbox{lightcyan}{8}{Len (Min: 0)} &
\bitbox{8}{[Data...]}
\end{bytefield}
	\begin{itemize}
		\item {\bf 0:} ACK. Indicates success.
		\item {\bf 1:} NAK. Generic error code indicating failure.
	\end{itemize}
	\item Unless otherwise specified...
	\begin{itemize}
		\item The remaining bytes shall be an ASCII-encoded string
		composed only of standard printable characters. The string
		shall not be NUL-terminated. The contents of this string is
		not well-defined and is expected to be something
		human-readable and useful.
	\end{itemize}
	\item If specified...
	\begin{itemize}
		\item The response is permitted to be implementation defined.
	\end{itemize}
\end{itemize}

\item \textbf{\texttt{0x56 (`V')} -- Query Versions}
\item {\em Synchronous Request}
\begin{itemize}
	\item[]
\begin{bytefield}{24} \\
\colorbitbox{lightgreen}{8}{0x56} &
\colorbitbox{lightred}{8}{Event ID} &
\colorbitbox{lightcyan}{8}{Len (Must be 0)} &
\end{bytefield}
	\item This message queries the protocol version(s) this ICE
implementation understands.
	\item If multiple versions are supported, they should be listed in
order of version preference.
	\item Response:

		\begin{bytefield}{56} \\
		\colorbitbox{lightgreen}{8}{0x00} &
		\colorbitbox{lightred}{8}{Event ID} &
		\colorbitbox{lightcyan}{8}{Len (Min: 2)} &
		\bitbox{8}{Major} &
		\bitbox{8}{Minor} &
		\bitbox{8}{[Major...]} &
		\bitbox{8}{[Minor...]} &
		\end{bytefield}
\end{itemize}

\item \textbf{\texttt{0x76 (`v')} -- Request to use version}
\item {\em Synchronous Request}
\begin{itemize}
	\item[]
\begin{bytefield}{40} \\
\colorbitbox{lightgreen}{8}{0x76} &
\colorbitbox{lightred}{8}{Event ID} &
\colorbitbox{lightcyan}{8}{Len (Must be 2)} &
\bitbox{8}{Major} & \bitbox{8}{Minor}
\end{bytefield}
	\item The message shall be composed of exactly two bytes.
	\item Each byte shall be an unsigned number.
	\item The first byte shall be considered a major version number.
	\item The second byte shall be considered a minor version number.
	\item Response:
	\begin{itemize}
		\item[]
\begin{bytefield}{24} \\
\colorbitbox{lightgreen}{8}{0x00} &
\colorbitbox{lightred}{8}{Event ID} &
\colorbitbox{lightcyan}{8}{Len (Must be 0)}
\end{bytefield}
		\item If the selected version is acceptable, an ACK shall be generated.
		\item[]
\begin{bytefield}{40} \\
\colorbitbox{lightgreen}{8}{0x01} &
\colorbitbox{lightred}{8}{Event ID} &
\colorbitbox{lightcyan}{8}{Length} &
\bitbox{8}{[Major]} & \bitbox{8}{[Minor]} &
\end{bytefield}
		\item Otherwise a NAK shall be generated.
		\item The NAK may optionally include a preferred version or
list of versions in the same format as the `V' response.
	\end{itemize}
	\item This message is not queriable.
\end{itemize}

\item \textbf{\texttt{0x58 (`X')} -- eXtension}
\item \textbf{\texttt{0x78 (`x')} -- extension}
\begin{itemize}
\item These characters are reserved for future e{\bf X}tentions.
\item No further specification is defined here.
\end{itemize}

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagebreak
\section{Protocols}
\label{sec:protocols}

The following protocols are currently well-defined:
\setcounter{tocdepth}{2}
%\etocsettocstyle{\subsection*{Versions}\hrule\medskip
%\begin{minipage}{.95\linewidth}}
%{\end{minipage}\medskip\hrule}
\etocsettocstyle
    {\subsection*{\contentsname}\hrule\medskip
        \everypar{\rightskip.1\linewidth}}
    {\nobreak\medskip\hrule\bigskip}
\localtableofcontents

\clearpage

\input{protocol-0-1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagebreak
\section{Document Revision History}
\label{sec:revisions}

\begin{itemize}

\item Revision 0.2.5 -- Jan 14, 2013
\subitem Add `p' and `P' messages

\item Revision 0.2.4 -- Dec 20, 2012
\subitem Add `g' and `G' messages

\item Revision 0.2.3 -- Dec 18, 2012
\subitem Add `f', `o', and `O' messages

\item Revision 0.2.2 -- Dec 15, 2012
\subitem `d' fragments are 255 bytes long so that a final fragment of maximum
length can be distinguished
\subitem Make explicit that each `d' fragment message is ACK/NAK'd
\subitem When a `d' fragment is NAK'd, that ends the transaction

\item Revision 0.2.1 -- Nov 29, 2012
\subitem 0.1's `d' $\rightarrow$ `e'
\subitem Add fragmented `d' messages

\item Revision 0.2 -- Nov 27, 2012
\subitem Largely redefine the protocol (ish)

\item Revision 0.1.2 -- Nov 26, 2012
\subitem Restrict Version to `V' only
\subitem Restrict eXtension to `X' only

\item Revision 0.1.1 -- Nov 16, 2012
\subitem Add visualized packets

\item Revision 0.1 -- Nov 14, 2012
\subitem Initial revision

\end{itemize}

\end{document}
