\documentclass{article}
%\usepackage{fullpage}
\usepackage[top=1in, bottom=1in, left=1cm,right=1cm]{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}

\begin{document}

\fancyfoot[L]{Revision 1.0 -- 11/17/2011}
\pagestyle{fancyplain}

\title{EECS 373 Cortex M3 Simulator Cleanup}
\author{Date Open: Thu Nov 17}
\date{{\bf Final submissions: 5PM, Mon Dec 12}}
\maketitle

One key aspect of good software engineering is called {\em regression
testing}. Unfortunately, we neglected to emphasize this point throughout the
simulator project. As a result, test cases that once worked became broken when
subsequent bugfixes or new features either created or exposed other bugs.

We would like to clean up the simulator and bring it to a completely working
state. However, we would not like to detract from your final course project,
as that should be your focus at this time.

The compromise then is the carrot not the stick. We are offering {\tt 0.5\%}
of extra credit for each failing test case that you fix. You may fix a maximum
of {\bf 5} test cases (for credit). This extra credit will be applied at the
very end of final grade calculation, after any curving or other corrections
are made to the overall distribution; it will not in any way affect students
who do not earn this extra credit.

\section{The Gory Details}
\subsection {Fixing Bugs}
In order to receive credit, you must explain {\em how} you fixed the problem.
In addition, you must {\em actually fix} the problem, not simply remove it,
e.g.:
\begin{itemize}
\item {\bf BAD:} The {\tt and\_testbench} was failing on the test case in
lines 78-92, so I removed them.
\item {\bf GOOD:} In {\tt programs/tests/log\_ext\_shift/and\_testbench} the
test case in lines 78-92 exposed a bug where the AND Immediate T2 encoding did
not properly set the status flags (lines 589-622 in {\tt
cpu/operations/logical.c}).  I corrected the instruction implementation to set
the flags correctly and the test case now passes.
\end{itemize}
As a general rule, if faced with two options of how to fix a bug you've
identified, the harder option is correct (this is extra credit after all).

\subsection{Claiming Credit}
Since it is possible that multiple people may be working on the same bugs in
parallel, we will rely on the subversion repository to break any ties. The
rule is simple: First to commit a fix wins.

\vspace{1 em}
{\bf One CRITICAL caveat:} Since the whole point of this is to address
regressions, introducing a new regression with your fix would be a
ReallyBadIdea\texttrademark, and would certainly invalidate your ``fix''.
\vspace{1 em}

The commit message for your fix should indicate clearly the bug you have fixed
and any test cases that now pass as a result of your bugfix. It is possible
that one bug fixes multiple test cases. The extra credit is awarded based on
the number of {\em test cases} fixed. We believe it appropriate to reward
those ``killer'' bugs that break a large number of test cases, rather than a
more subjective ``per bug'' style system.

\subsection{The Testcases}
As a starting reference, the current status of all testcases have been
appended to the end of this document. You can use the supplied {\tt
run\_tests.sh} to run them in batch, or simply run them each by hand.
\clearpage
\begin{verbatim}
                                programs/basic.bin: PASSED
                                programs/blink.bin: FAILED
                                 programs/echo.bin: FAILED
                             programs/echo_str.bin: FAILED
                       programs/tests/trivialC.bin: PASSED
                  programs/tests/trivialPrintf.bin: FAILED
                       programs/tests/trivialS.bin: PASSED
      programs/tests/branch/cbnz_and_cbz_combo.bin: PASSED
               programs/tests/branch/cbnz_test.bin: PASSED
                programs/tests/branch/cbz_test.bin: PASSED
           programs/tests/branch/cmp_testbench.bin: FAILED
      programs/tests/add_sub_mov/add_testbench.bin: FAILED
      programs/tests/add_sub_mov/mov_testbench.bin: FAILED
      programs/tests/add_sub_mov/sub_testbench.bin: FAILED
    programs/tests/log_ext_shift/and_testbench.bin: PASSED
    programs/tests/log_ext_shift/bic_testbench.bin: PASSED
    programs/tests/log_ext_shift/eor_testbench.bin: FAILED
 programs/tests/log_ext_shift/extend_testbench.bin: PASSED
    programs/tests/log_ext_shift/mvn_testbench.bin: FAILED
    programs/tests/log_ext_shift/orn_testbench.bin: PASSED
    programs/tests/log_ext_shift/orr_testbench.bin: PASSED
  programs/tests/log_ext_shift/shift_testbench.bin: FAILED
    programs/tests/log_ext_shift/teq_testbench.bin: FAILED
    programs/tests/log_ext_shift/tst_testbench.bin: FAILED
               programs/tests/ldr_str/ldrb_imm.bin: PASSED
               programs/tests/ldr_str/ldrb_lit.bin: FAILED
               programs/tests/ldr_str/ldrb_reg.bin: FAILED
     programs/tests/ldr_str/ldrd-imm_testbench.bin: FAILED
     programs/tests/ldr_str/ldrd-lit_testbench.bin: FAILED
     programs/tests/ldr_str/ldrh_imm_testbench.bin: FAILED
     programs/tests/ldr_str/ldrh_lit_testbench.bin: PASSED
     programs/tests/ldr_str/ldrh-reg_testbench.bin: PASSED
      programs/tests/ldr_str/ldr_lit_testbench.bin: FAILED
              programs/tests/ldr_str/ldrsb_imm.bin: FAILED
              programs/tests/ldr_str/ldrsb_lit.bin: FAILED
              programs/tests/ldr_str/ldrsb_reg.bin: FAILED
    programs/tests/ldr_str/ldrsh-imm_testbench.bin: PASSED
    programs/tests/ldr_str/ldrsh_lit_testbench.bin: PASSED
  programs/tests/ldr_str/ldrstrh_reg_testbench.bin: FAILED
   programs/tests/ldr_str/ldrstr_imm_testbench.bin: FAILED
   programs/tests/ldr_str/ldrstr_reg_testbench.bin: FAILED
 programs/tests/ldr_str/ldrstrsh_reg_testbench.bin: FAILED
               programs/tests/ldr_str/strb_imm.bin: FAILED
               programs/tests/ldr_str/strb_reg.bin: FAILED
     programs/tests/ldr_str/strd-imm_testbench.bin: FAILED
     programs/tests/ldr_str/strh_imm_testbench.bin: FAILED
     programs/tests/ldr_str/strh-reg_testbench.bin: PASSED
         programs/tests/push_pop/pop_testbench.bin: PASSED
        programs/tests/push_pop/push_testbench.bin: FAILED
          programs/tests/mul_div/div_testbench.bin: PASSED
         programs/tests/mul_div/mul_tb_shafeen.bin: FAILED
      programs/tests/mul_div/mul_testbench_fim.bin: PASSED
\end{verbatim}

\end{document}
